---
title: 'Bios 6301: Homework 7'
author: Ashley Spirrison 
output: pdf_document
---

*Due Thursday, 02 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

40 points total.

Submit a single knitr file (named `homework7.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework7.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**21 points**

Use the following code to generate data for patients with repeated measures of A1C (a test for levels of blood glucose).

```{r}
genData <- function(n) {
    if(exists(".Random.seed", envir = .GlobalEnv)) {
        save.seed <- get(".Random.seed", envir= .GlobalEnv)
        on.exit(assign(".Random.seed", save.seed, envir = .GlobalEnv))
    } else {
        on.exit(rm(".Random.seed", envir = .GlobalEnv))
    }
    set.seed(n)
    subj <- ceiling(n / 10)
    id <- sample(subj, n, replace=TRUE)
    times <- as.integer(difftime(as.POSIXct("2005-01-01"), as.POSIXct("2000-01-01"), units='secs'))
    dt <- as.POSIXct(sample(times, n), origin='2000-01-01')
    mu <- runif(subj, 4, 10)
    a1c <- unsplit(mapply(rnorm, tabulate(id), mu, SIMPLIFY=FALSE), id)
    data.frame(id, dt, a1c)
}
x <- genData(500)
```



```{r}
#2. For each `id`, determining if there is more than a one year gap in between observations. Adding a new row at the one year mark (two for two years, etc.), with the `a1c` value set to missing. 

addMissingRows <- function(data) {
  result <- vector("list", length = nrow(data))
  k <- 1

  data <- data[order(data$id, data$dt), ]  #1. Ordering the data set by `id` and `dt`.

  for (i in unique(data$id)) {
    temp_data <- data[data$id == i, ]
    temp_data$dt <- as.POSIXct(temp_data$dt)  
    temp_data <- temp_data[order(temp_data$dt), ]  

    for (j in 1:(nrow(temp_data) - 1)) {
      time_diff <- as.numeric(difftime(temp_data$dt[j + 1], temp_data$dt[j], units = "days"))
      if (!is.na(time_diff) && time_diff >= 365) {
        num_missing_years <- floor(time_diff / 365)
        for (m in 1:(num_missing_years - 1)) {
          if (m == 1) {
            new_dt <- as.POSIXct(temp_data$dt[j] + as.difftime(365 * m, units = "days"))  
            new_row <- data.frame(id = i, dt = new_dt, a1c = NA)
            temp_data <- rbind(temp_data[1:j, ], new_row, temp_data[(j + 1):nrow(temp_data), ]) 
            j <- j + 1  
          }
        }
      }
    }
    result[[k]] <- temp_data
    k <- k + 1
  }

  result <- do.call(rbind, result)
  return(result)
}

x <- addMissingRows(x)

#3. Creating a new column `visit`.  For each `id`, adding the visit number from 1 to n depending on the number of observations for each individual. Observations created with missing a1c values are included.
x$visit <- ave(x$id, x$id, FUN = seq_along)

#4. For each `id`, determining and replacing missing values with the mean `a1c` value for each individual.
mean_a1c <- aggregate(a1c ~ id, data = x, FUN = function(x) mean(x, na.rm = TRUE))

for (i in 1:nrow(mean_a1c)) {
  id_val <- mean_a1c[i, "id"]
  mean_val <- mean_a1c[i, "a1c"]
  x$a1c[x$id == id_val & is.na(x$a1c)] <- mean_val
}

#5.Printing the mean a1c for each id.
print(mean_a1c)

#6.Calculating and printing the total number of visits for each id.
total_visits <- aggregate(visit ~ id, data = x, FUN = length)

print(total_visits)

#7.Printing the observations for id = 15.
print(x[x$id == 15, ])



```



### Question 2 ###

**16 points**

Install the `lexicon` package. Load the `sw_fry_1000` vector, which contains 1,000 common words.

```{r}
options(repos = c(CRAN = "http://cran.rstudio.com/"))


install.packages("lexicon")
library(lexicon)
data('sw_fry_1000', package = 'lexicon')
head(sw_fry_1000)
```

```{r}
#1. Removing all non-alphabetical characters and making all characters lowercase. Saving the result as a. 

a <- tolower(gsub("[^[:alpha:]]", "", sw_fry_1000))
save(a, file = "a.RData")
print(a)
```



```{r}
#2. Determining the number of words containing the string "ar". 
count <- sum(grepl("ar", a))
print(count)

```

```{r}
#3. Finding a six-letter word that starts with "l" and ends with "r".  
six_letter_word <- a[grep("^l[a-z]{4}r$", a)]
print(six_letter_word)
```


```{r}
#4. Returning all words starting with "col" or end with "eck".
result <- a[grep("^col|eck$", a)]
print(result)

```

```{r}
#5. Finding the number of words that contain 4 or more consonants (with y included as a consonant).
pattern <- "(.*[bcdfghjklmnpqrstvwxyz]{4,}.*)"
count <- sum(grepl(pattern, a))
print(count)

```

```{r}
#6 Returning all words that have a "q" not followed by a "ui".
result <- a[grep("q(?!ui)", a, perl = TRUE)]
print(result)

```
```{r}
#7. Finding all words that contain "k" followed by another letter. Running table command on first character following first "k" of each word. 
matches <- grep("k[A-Za-z]", a, value = TRUE)

following_chars <- regmatches(matches, regexpr("(?<=k).", matches, perl = TRUE))

char_table <- table(following_chars)

print(char_table)

```



```{r}
#Removing all vowels and determining the number of character strings found once. 
a_no_vowels <- gsub("[aeiou]", "", a)

unique_count <- length(which(table(a_no_vowels) == 1))

print(unique_count)

```
**3 points**

The first argument to most functions that fit linear models are formulas.  The following example defines the response variable `death` and allows the model to incorporate all other variables as terms. `.` is used to mean all columns not otherwise in the formula.

```{r}
url <- "https://github.com/couthcommander/Bios6301/raw/main/datasets/haart.csv"
haart_df <- read.csv(url)[,c('death','weight','hemoglobin','cd4baseline')]
coef(summary(glm(death ~ ., data=haart_df, family=binomial(logit))))
```

Now imagine running the above several times, but with a different response and data set each time.  Here's a function:

```{r}
myfun <- function(dat, response) {
  form <- as.formula(response ~ .)
  coef(summary(glm(form, data=dat, family=binomial(logit))))
}
```

Unfortunately, it doesn't work. `tryCatch` is "catching" the error so that this file can be knit to PDF.

```{r}
tryCatch(myfun(haart_df, death), error = function(e) e)
```

```{r}
#1.  Debugging the myfun function
debug(myfun)
tryCatch(myfun(haart_df, "death"), error = function(e) e)
#through debugging, an error message is recieved that indicates that variable lengths differ. I believe this is due to the as.formula(response~.) section, which indicates that a variable 'response' should appear as a column in dat. The formula construction uses a character string and therefore does not recognize the variable 'response' as a column of dat. The 'variable lengths differ' error message is due to the function not being able to find the response variable in the dataset. 



```

```{r}
#bonus. creating a working function.

url <- "https://github.com/couthcommander/Bios6301/raw/main/datasets/haart.csv"
haart_df <- read.csv(url)[,c('death','weight','hemoglobin','cd4baseline')]
coef(summary(glm(death ~ ., data=haart_df, family=binomial(logit))))

myfun <- function(dat, response) {
  form <- as.formula(paste(response, "~ ."))
  coef(summary(glm(form, data=dat, family=binomial(logit))))
}
# Creating a sample dataset
set.seed(123)  
n <- 100
dat <- data.frame(
  response = sample(c(0, 1), n, replace = TRUE),
  x1 = rnorm(n),
  x2 = rnorm(n),
  x3 = rnorm(n)
)

# Testing the function
result <- myfun(dat, "response")
print(result)


```
