---
title: "homework2"
author: "Ashley Spirrison"
date: "2023-10-01"
output: github_document
---

# 1. Loading cancer data set into R and making it a data frame called cancer.df.
```{r}
    file <- 'cancer.csv'
    data <- data.frame(read.csv(file))
    cancer.df <- data.frame(data)
    cancer.df
    getwd()
    
```
## Determining the number of rows and columns in the data frame. 
    ```{r}
    nrow(cancer.df)
    ncol(cancer.df)
    ```
## Extracting the names of the columns in cancer.df. 
    
    ```{r}
    colnames(cancer.df)
    ```

## Reporting the value of the 3000th row in column 6.
    ```{r}
    cancer.df[3000,6]
    ```

## Reporting the contents of the 172nd row. 
    ```{r}
    cancer.df[172,]
    ```

## Creating a new column that is the incidence *rate* (per 100,000) for each row. 
    ```{r}
    cancer.df[,'new']<-(cancer.df$incidence/cancer.df$population)*100000
    cancer.df
    ```
## Determining how many subgroups (rows) have a zero incidence rate.
    ```{r}
    nrow(cancer.df[cancer.df$new == 'zero'])
    ```

## Finding the subgroup with the highest incidence rate.
    ```{r}
    cancer.df[which.max(cancer.df$new),]
    ```

# 2. Data types.

## Creating the following vector: x <- c("5","12","7"). Determining which of the following commands will produce an error message. For each command, Either explaining why they should be errors, or explaining the non-erroneous result. 
```{r}
            x <- c("5","12","7")
            max(x) 
            #no error, max function is case-insensitive, can use characters 
            sort(x)
            #sort can be utilized to for characters, however the result seems a bit odd because with characters sort tries to put items in ascending order, it looks at the 1 first even though 12 is greater than 5 and 7
            #sum(x)
            #there is an error here because you cannot sum (or any other numerical operation) a characters, you would need to instead make the vector numerics then you could find the sum 
            class(x) 
```

## For the next two commands, either explaining their results, or why they should produce errors. 

```{r}
            #y <- c("5",7,12)
            #y[2] + y[3]
            #class(y)
            #there is an error here because y contains characters and you cannot perform binary operations on non-numerics, if you remove the "5" you do not recieve an error (see below)
            #fix <-c(5,7,12)
            #fix[2] + fix[3]
```
            

## For the next two commands, either explaining their results, or why they should produce errors. 
```{r}
            z <- data.frame(z1="5",z2=7,z3=12)
            z
            z[1,2] + z[1,3]
            class(z)
            #there is not an error here because instead of being a vector, z is now a dataframe. You are selecting for z2 and z3, which are both numerics and can perform the addition without problem.
```

# 3. Data structures.
## Given R expressions and returning matrices and vectors. 

## $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
```{r}  
    t <- seq(1,8)
    l <- seq(7,1)
    con <-c(t,l)
    con
```
    
## $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
```{r}
    
    rep(1:5, times = 1:5)
```

## $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$
    
```{r}
    row <- 3
    col <- 3
    diag_mat <- matrix(1, row, col)
    diag(diag_mat) <- 0
    print(diag_mat)
```
    
## $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$
```{r}
    nm <- cbind(rep(1, 4),2^(1:5),3^(1:5),4^(1:5))
    nm
```
    
    
# 4. Basic programming.

## Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. As an example, use `x = 5` and `n = 2`. 
```{r}
   x = 5
   n = 2
   h = 0
   for (i in 1:n)
   {current = x^i
   h = h + current}
   h
```
    


## Finding the sum of all the multiples of 3 or 5 below 1,000. 
```{r}
       x <- c(seq(0,999,5))
       x
       y <- c(seq(0,999,3))
       y
       z <- sum(x,y)
       z
```
       
## Finding the sum of all the multiples of 4 or 7 below 1,000,000. 
```{r}
       x <- c(seq(0,999999,4))
       x
       y <- c(seq(0,999999,7))
       y
       z <- sum(x,y)
       z
```
## Writing an R program to calculate the sum of the first 15 even-valued terms.
```{r}
y=c(1,2,3,5,8,13,21,34,55,89, 144, 233, 377, 610, 987, 1597, 2584, 4181, 6765, 10946, 17711, 28657, 46368, 75025, 121393, 196418, 317811, 514229, 832040, 1346269, 2178309, 3524578, 5702887, 9227465, 14930352, 24157817, 39088169, 632245986, 102334155, 165580141, 267914296, 433494437, 701408733, 1134903170)
y
y_logical <- y %% 2 == 0
y_logical 
y_even <- y[y_logical]
y_even
z=sum(y_even)
z

```

