---
title: 'Bios 6301: Assignment 5'
author: Ashley Spirrison 
output: pdf_document
---

*Due Thursday, 12 October, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

40 points total.

Submit a single knitr file (named `homework5.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework5.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

```{r}

#writing secant function, validating, and finding iterations to convergence 
secant_function <- function(x) { 
  return(cos(x)-x)}
x0 <- 0
x1 <- 1
tolerance <- 1e-9
max_iterations <- 100
iterations <- 0
while(iterations < max_iterations) {
  x2 <- x1 -(secant_function(x1) * (x1-x0)) / (secant_function(x1) - secant_function(x0))
  if (abs(x2-x1) < tolerance) {
    cat("converged to root: ", x2, "\n")
    return(x2)
    break
  }
  x0 <- x1
  x1 <- x2
  iterations <- iterations + 1
}
iterations

#writing N-R function and finding iterations to convergence to allow for secant comparison 

n_r_function <-function(x) {
  return (cos(x) -x)
}
n_r_deriv <- function(x) {
  return(-sin(x) - 1)
}
x0 <- 0
tolerance <- 1e-9
max_iterations <- 100
iterations <- 0
while (iterations < max_iterations) {
  x1 <- x0 -n_r_function(x0)/ n_r_deriv(x0)
  if(abs(x1-x0) < tolerance) {
  cat("converged to root (N-R):", x1, "\n")
  return(x1)
  break
}
x0 <- x1
iterations <- iterations + 1
}
iterations

#N-R method is faster, coverged to root in 4 iterations versus 5 iterations for secant method

```

### Question 2 ###

**20 points**


```{r}
  #using verbose to control output 
#creating craps function 
craps <- function(verbose = TRUE) {
  x <- sum(ceiling(6 * runif(2)))
  if (verbose) cat("First roll:", x, "\n")
  
  if (x %in% c(7, 11)) {
    if (verbose) cat("win\n")
    return(TRUE)
  } else {
    y <- x
    if (verbose) cat("y is", y, "\n")
    repeat {
      roll <- sum(ceiling(6 * runif(2)))
      if (verbose) cat("Next roll:", roll, "\n")
      if (roll == y) {
        if (verbose) cat("win\n")
        return(TRUE)
        
      } else if (roll == 7 || roll == 11) {
        if (verbose) cat("lose\n")
        return(FALSE)
        
      }
    }
  }
}

#setting seed and running 3 games 
set.seed(100)

for (i in 1:3) {
  craps()
}

#determining seed to win 10 straight games 
deter_w_seed <- function() {
  seed <- 1
  while (TRUE) {
    set.seed(seed)
    wins <- 0
    consec_wins <- 0
    while (consec_wins < 10) {
      if (craps(FALSE)) {
        consec_wins <- consec_wins + 1
      } else {
        consec_wins <- 0
      }
    }
    if (consec_wins == 10) {
      cat("Seed", seed, "results in ten consecutive wins\n")
      for (i in 1:10) {
        cat("Game", i, "Result: win\n\n")
      }
      break
    }
    seed <- seed + 1
  }
}

deter_w_seed()


```


### Question 3 ###

**5 points**

This code makes a list of all functions in the base package:

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
```

Using this list, write code to answer these questions.

1. Which function has the most arguments? (3 points)
```{r}
most_argue <- 0
most_argue_function <- NULL 
for(fun in funs) {
  argue <- length(formals(fun))
  if (argue > most_argue) {
    most_argue <- argue
    most_argue_function <- fun
  }
}

cat("function with most arguements:", deparse(most_argue_function), "with", most_argue, "arguements.\n")

```

1. How many functions have no arguments? (2 points)

```{r}
no_argue_functions <- 0
for(fun in funs) {
  argue <- length(formals(fun))
  if (argue == 0) {
    no_argue_functions <- no_argue_functions +1
    
  }
}

cat("number of functions with no arguements:", no_argue_functions, "\n")

```
Hint: find a function that returns the arguments for a given function.
