---
title: "homework1"
author: "Ashley Spirrison"
date: "2023-10-01"
output: github_document
---

### Creating the Data Set


```{r}
gender <- c('M','M','F','M','F','F','M','F','M')
age <- c(34, 64, 38, 63, 40, 73, 27, 51, 47)
smoker <- c('no','yes','no','no','yes','no','no','no','yes')
exercise <- factor(c('moderate','frequent','some','some','moderate','none','none','moderate','moderate'),
                    levels=c('none','some','moderate','frequent'), ordered=TRUE
)
los <- c(4,8,1,10,6,3,9,4,8)
x <- data.frame(gender, age, smoker, exercise, los)
x
```

### Creating the Model


```{r}
lm(los ~ gender + age + smoker + exercise, dat=x)
```

### Coefficient has the highest effect on 'los' is genderM



### Creating a model using los and gender and assign it to the variable mod and running the summary function with mod as its argument.

```{r}
gender <- c('M','M','F','M','F','F','M','F','M')
los <- c(4,8,1,10,6,3,9,4,8)
mod<-lm(los ~ gender,dat=x)
summary(mod)
```



### Determining the estimate for the intercept and gender using coef. funct.

```{r}
coef(mod)
#intercept: 3.5, genderM: 4.3
``` 


### Calculating the standard errors. 

```{r}
sqrt(diag(vcov(summary(mod))))

```

### Calculating test statistics.

```{r}
mod <- lm(los ~ gender, dat=x)
mod.c <- coef(summary(mod))
mod.c[,1]/mod.c[,2]

```


### Calculating the p value for gender.

```{r}
pt(mod.c[,1]/mod.c[,2],7,lower.tail=FALSE)*2
#gender p-value:0.0224

```

### Predicted Values generation.


```{r}
3.5+(x$gender=='M')*4.3
```

### Passing model to predict and fitted.

```{r}
predict(mod)
fitted(mod)
```

### Passing newdat to predict.

```{r}
newdat <- data.frame(gender=c('F','M','F'))
predict(mod,newdat)
```

### Residuals determination.

```{r}
x$los-predict(mod)
```

### Passing mod to residuals.

```{r}
residuals(mod)
```

# Squaring residuals, summing, and comparing to passing mod to def.

```{r}
t<-residuals(mod)
sum(t^2)
deviance(mod)
#result in equivalent values
```

### Passing mod to df. residual.
```{r}
df.residual(mod)
```

### Calculating SE.

```{r}
y<-deviance(mod)/df.residual(mod)
sqrt(y)
#standard error:2.197
```

###Noting it matches the output of the below line.

```{r}
predict(mod, se.fit=TRUE)$residual.scale
```

### Running t-test. 

## Creating subset of x where gender is M and assigning to men, doing same with women.

```{r}
men<-subset(x,gender=='M')

women<-subset(x,gender=='F')

```

### Calculating variance.

```{r}
#men variance for los
var(men$los)
#women variance for los
var(women$los)
#variance: men:5.2, women:4.33
```

### Calling t-test function.

```{r}
t.test(women$los,men$los)
t.test(women$los,men$los,var.equal=TRUE)
#the second one (where var.equal=TRUE) matches the p value for gender from the model summary.

```

### Alternative way to calling t-test.

```{r}
t.test(los ~ gender, dat=x, var.equal=TRUE)
# compare p-values
t.test(los ~ gender, dat=x, var.equal=TRUE)$p.value
coef(summary(lm(los ~ gender, dat=x)))[2,4]
```
